---
title: '<font size = 7 color = "black">Multiple Linear Regression </font>'
subtitle: '<font size = 6 color = "black"> Model Building and Analysis of Flight Delay Times </font>'
author: '<font size = 5 color = "black">Ian Vanwright </font>'
institute: '<font size = 6 color = "white">West Chester University of Pennsylvania</font><br> '
date: ' <font color = "black" size =4> Prepared for<br> </font> <font color = "gold" size = 6> STA490: Data Visualization </font> <br><br> <font color = "white" size = 3></font>'
output:
  xaringan::moon_reader:
    css: xaringan-themer01.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("pander")) {
   install.packages("pander")
   library(pander)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}
if (!require("kableExtra")) {
  install.packages("kableExtra")
  library(kableExtra)
}
knitr::opts_chunk$set(
                  fig.width=3, 
                  fig.height=3, 
                  fig.retina=12,
                  out.width = "100%",
                  cache = FALSE,
                  echo = FALSE,
                  message = FALSE, 
                  warning = FALSE,
                  hiline = TRUE
                  )
```


```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
  style_duo_accent(primary_color = "#1F4257",
          secondary_color = "#380F2A",
          # fonts
          header_font_google = google_font("Martel"),
          text_font_google = google_font("Lato"),
          code_font_google = google_font("Fira Mono"))
  

```


```{r, include=FALSE}
 if (!require("xaringanthemer")) {
      install.packages("xaringanthemer")
      library(xaringanthemer)
     }
```

```{r}
flight1<- read.csv("flight_delay-data.csv")
flight2 <- flight1[-which(flight1$Arr_Delay == 0),]
```

class: top, center
#  Table of Contents: 

.left[

- Introduction 

- Research Question

- Recap of Building the Final Model

- Bootstrapping Part 1: Observations 

- Bootstrapping Part 2: Residuals

- Summary of Results/Model Comparisons

- Conclusion
]


---

class: top, center
# Introduction: 


.left[


- Data Set was taken from 'Applied Analytics Through Case Studies Using SAS and R by Deepti Gupta'

- Data set consists of 3593 observations and 11 variables

- Variables are as follows:
   - 1) Carrier (Categorical) - The airline that the flight is taken with
   - 2) Airport_Distance (Numeric) - The Distance between the airports (in miles)
   - 3) Number_of_flights (Numeric) - Total number of flights in airport
   - 4) Weather (Numeric) - Delay due to weather condition ranked 0-10, with 0 being mild and 10 being extreme 
   - 5) Support_Crew_Available (Numeric) - Total number of support crew available
   - 6) Baggage_loading_time (Numeric) - Time (in minutes) for loading the baggage
  - 7) Late_Arrival_o (Numeric) - Time (in minutes) for late arriving aircraft of the same flight
  - 8) Cleaning_o (Numeric) - Time (in minutes) for aircraft cleaning
  - 9) Fueling_o (Numeric) - Time (in minutes) for aircraft fueling
  - 10) Security_o (Numeric) - Time (in minutes) for security checking
  - 11) Arr_Delay (Numeric) -  Flight arrival time (in minutes). This will be the dependent variable in the model

]

---

class: top, center
# Summary of Building the Final Model

.left[

- Fit the full model with Arr_Delay as the response and all other variables as predictors

- T-tests determined Cleaning_o, Fueling_o, Security_o were all insignificant (p > 0.05) and I decided to remove carrier (this created the reduced model, which analysis was done with)

- Diagnostics of the reduced model were checked, and the assumptions of normality and homoscedasticity (constant variances) were violated

- A box-cox transformation was applied to Arr_Delay  (λ = 0.5) , fixing issues with normality and constant variances

- Comparing the goodness-of-fit measures of the full model, the reduced model, and the model with the transformation, the model with the transformation was chosen


]
---

class: top, center
# Final Model

.center[

<div style="text-align: center;">
  <img src= "img/MLRInitialFinalModel.PNG" width = 600 height=450;>
</div>


]

---

class: top, center
# Residual Analysis of the Final Model

<div style="text-align: center;">
  <img src= "img/MLRResidualAnalysis.PNG" width = 300 height=300;">
  <img src= "img/MLRResidualHistogram.PNG" width = 300 height=300;">
  <img src= "img/MLRMeanResiduals.PNG" width = 300 height=300;">
</div>


---
class: top, center
# Bootstrapping (Part 1): Observations

.left[

- Bootstrapping: Resampling with replacement

- Because our sample size is big enough, bootstrapping is a good method to use

- Sampling Distributions of Coefficient Estimates are pictured 

- Red density curves represent the actual sampling distribution of the coefficients, while the blue density curves represent the sampling distribution created from bootstrapping 

]

<div style="text-align: center;">
  <img src= "img/MLRBootstrap1.PNG" width = 300 height=300;">
  <img src= "img/MLRBootstrap2.PNG" width = 300 height=300;">
  
</div>

---
class: top, center
# Bootstrapping Part 1 (Continued)

- We can check how well the bootstrap did by summarizing the results and building a confidence interval

- The estimated parameters and the 95% confidence interval for the bootstrapped sampling distribution are pictured 

- All parameter estimates in the coefficient matrices match (within rounding)

<div style="text-align: center;">

  <img src= "img/MLRInitialFinalModel.PNG" width = 300 height=300;">
  <img src= "img/MLRBootstrapModel1.PNG" width = 300 height=300;">

</div>

---

class: top, center
# Bootstrapping (Part 2): Residuals

.left[

- Now we will look at the residuals of our sample and resample 

- Residual sampling distribution values: fitted value + residual

- Sampling distributions of coefficient estimates are pictured
Red density curves represent the actual sampling distribution of the coefficients, while the blue density curves represent the sampling distribution created from bootstrapping 

]

<div style="text-align: center;">
  <img src= "img/MLRBootstrap3.PNG" width = 300 height=300;">
  <img src= "img/MLRBootstrap4.PNG" width = 300 height=300;">

</div>

---

class: top, center
# Boostrap Part 2 (Continued)

- Again, we can check how well the bootstrap did by summarizing the results and building a confidence interval

- The estimated parameters and the 95% confidence interval for the bootstrapped residual sampling distribution are pictured

- Again, all parameter estimates in the coefficient matrices match (within rounding)

<div style="text-align: center;">
  <img src= "img/MLRInitialFinalModel.PNG" width = 300 height=300;">
  <img src= "img/MLRBootstrapModel2.PNG" width = 300 height=300;">

</div>

---

class: top, center
# Summary of Results

- Essentially no difference in parameter estimation for all 3 models

- Will choose parametric (original) model 

<div style="text-align: center;">
  <img src= "img/MLRInitialFinalModel.PNG" width = 400 height=400;">
  

</div>

---
class: middle, center
# That's all, folks!

# Any Questions, Comments, Thoughts, Concerns?

